---
title: "pre_lab_04.Rmd"
author: "derek willis"
date: "8/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Points to hit
1. Review of third lab questions/problems.
2. Demonstration of combining and merging

## Chapter 10

### Task 1: Load libraries and settings
**Task** Run the following code in the gray-colored codeblock below to load the tidyverse library and turn off scientific notation.  

```{r}
# turn off sci notation
options(scipen=999)
library(tidyverse)
library(lubridate)
```

### Task 2: Load data
Let's say that we have county population estimates for three different years - 2010, 2015 and 2020 - in three different files. They have the same record layout and the same number of counties.

**Task** Load the three CSV population estimate files. Do they have the same column structure?
**Answer** They do- each have 3194 rows and columns 9. 

```{r}
popestimate_2010 <- read_csv("data/popestimate_2010.csv")
```

```{r}
popestimate_2015 <- read_csv("data/popestimate_2015.csv")
```

```{r}
popestimate_2020 <- read_csv("data/popestimate_2020.csv")
```


### Task 3: Using bind_rows
**Task** Combine the three files into a single data frame using the bind_rows function.

```{r}
# bind_rows with list
estimates <- bind_rows(list(popestimate_2010, popestimate_2015, popestimate_2020))
View(estimates)
```

## Joining data

### Task 4: Loading data
**Task** Load the Maryland PPP data and a CSV file with NAICS codes. How many rows are in maryland_ppp?
**Answer** There are 195,869 rows in the maryland_ppp.

```{r}
maryland_ppp <- read_rds("pre_lab_04.rds")
```

```{r}
naics_codes <- read_csv('data/naics_codes.csv')
```

### Task 5: Join Them
**Task** Run the following code to join the PPP data to the NAICS codes using the common column naics_code as the key. How many rows are there now?
**Answer** There are still 195,869 rows in the maryland_ppp.

```{r}
# with nrow included to match row total
#start with larger data set, left join means show/connect the rows that are in common, select name (from md ppp df), naics_code (from both) and title (from naics code df), the count , nrow() means count all the rows
maryland_ppp %>% left_join(naics_codes, by="naics_code") %>% select(name, naics_code, title) %>% nrow()
```
### Task 6: Looking at Joined Data
**Task** Examine the combined data

```{r}
# without nrow
maryland_ppp %>% left_join(naics_codes) %>% select(name, naics_code, title)
```

### Task 7: Assign Joined Data to New Dataframe
**Task** Create a new dataframe with PPP and NAICS data joined together

```{r}
#creates new df that is the same ad md_ppp but adds a column for title of naics codes, if you filter or group by, etc. with this new df, title can be included in that
maryland_ppp_with_naics <- maryland_ppp %>% left_join(naics_codes)
```

### Task 8: Load the ZCTA Data
**Task** Load the Zip Code Tabulation Area data for Maryland

```{r}
maryland_zcta <- read_csv('data/maryland_zcta.csv')
```

### Task 9: Join ZCTA to PPP Data
**Task** Join the ZCTA data to the our PPP & NAICS data. What's the error message you get? What do you think it means?
**Answer** It's recognizing one of the zip columns as numbers, and the other column in the other file as a string of characters. 

```{r}
# join on zip - but there's a problem
#tell r that although they have diff column names, they mean the same thing/want to use them together
maryland_ppp_with_naics_and_demographics <- maryland_ppp_with_naics %>% left_join(maryland_zcta, by=c("zip"="ZCTA5N"))
```
### Task 10: Fix ZCTA Data
**Task** Fix the ZCTA file by changing the datatype

```{r}
#sees one as character and one as a number, so change it to a character
maryland_zcta <- maryland_zcta %>% mutate(across(ZCTA5N, as.character))
```

### Task 11: Join ZCTA to PPP Data and Check
**Task** Join the ZCTA data to the our PPP & NAICS data. Examine the columns in the new dataframe - does every row have demographic data? HINT: Use the >> character to browse columns
**Answer** All of the data is listed as NA.

```{r}
maryland_ppp_with_naics_and_demographics <- maryland_ppp_with_naics %>% left_join(maryland_zcta, by=c("zip"="ZCTA5N"))
View(maryland_ppp_with_naics_and_demographics)
```

### Task 12: Make a Zip5 Column, Join ZCTA to PPP Data and Check Again
**Task** Make a zip5 column and join again. Examine the new dataframe - now does every row have demographic data?
**Answer** Yes, now the demographic information is listened. 

```{r}
# make a zip5 column and redo the join
#make it a 5-digit zip instead of 9 digit zip, that's why it wouldn't originally join
maryland_ppp_with_naics <- maryland_ppp_with_naics %>% mutate(zip5 = str_sub(zip, 1, 5))
maryland_ppp_with_naics_and_demographics <- maryland_ppp_with_naics %>% left_join(maryland_zcta, by=c("zip5"="ZCTA5N"))
```

### Task 13: Asking Demographic Questions
**Task** Find the number of PPP loans that went to ZCTAs with more than 50% non-Hispanic Black population, and calculate the average and median amounts for those loans. Put your answers below.
**Answer** Number of PPP loans that went to ZCTAs with more than 50% non-Hispanic Black population: 44,032.	Average amount: 53836.46. Median amount: 20000.	

```{r}
# zcta with > 50% non-Hispanic Black population
maryland_ppp_with_naics_and_demographics %>%
  filter(PNHB > 50) %>%
  summarize(
    count = n(),
    avgamount = mean(amount),
    medamount = median(amount))
```


### Task 14: Asking Demographic Questions
**Task** Find the number of PPP loans that went to ZCTAs with more than 50% non-Hispanic white population, and calculate the average and median amounts for those loans. Put your answers below.
**Answer** Number of PPP loans that went to ZCTAs with more than 50% non-Hispanic white population: 113,937. Average amount: 88,140.24 . Median amount: 20,833.	

```{r}
# zcta with > 50% non-Hispanic white population
maryland_ppp_with_naics_and_demographics %>%
  filter(PNHW > 50) %>%
  summarize(
    count = n(),
    avgamount = mean(amount),
    medamount = median(amount))
```

### Task 15: Asking Demographic Questions
**Task** Find the number of PPP loans that went to ZCTAs with more than 50% non-Hispanic Black population, grouped by rural or urban status, and calculate the average and median amounts for those loans. Put your answers below.
**Answer** 
Rural: 
PPP loans that went to ZCTAs with more than 50% non-Hispanic Black population: 148
Average amount: 44,328.87	
Median Amount: 19,621	

Urban: 
PPP loans that went to ZCTAs with more than 50% non-Hispanic Black population: 43,884
Average amount: 53,868.53	
Median Amount: 20,000	

```{r}
# zcta with > 50% non-Hispanic Black population grouped by rural/urban
maryland_ppp_with_naics_and_demographics %>%
  filter(PNHB > 50) %>%
  group_by(rural_urban_indicator) %>%
  summarize(
    count = n(),
    avgamount = mean(amount),
    medamount = median(amount))
```

### Task 16: Asking Demographic Questions
**Task** Find the number of PPP loans that went to ZCTAs with more than 50% non-Hispanic white population, grouped by rural or urban status, and calculate the average and median amounts for those loans. Put your answers below.
**Answer**
Rural: 
PPP loans that went to ZCTAs with more than 50% non-Hispanic white population: 18,337	
Average amount: 73,694.14	
Median Amount: 20,833	

Urban: 
PPP loans that went to ZCTAs with more than 50% non-Hispanic white population: 95,600	
Average amount: 90,911.14	
Median Amount: 20,833	
```{r}
# zcta with > 50% non-Hispanic white population grouped by rural/urban
maryland_ppp_with_naics_and_demographics %>%
  filter(PNHW > 50) %>%
  group_by(rural_urban_indicator) %>%
  summarize(
    count = n(),
    avgamount = mean(amount),
    medamount = median(amount))
```
