---
title: "Reverse Engineering Project"
author: Jenna Cohen, Taneen Momeni, Max Boyarsky and Jamie Pinzon
date: 10/9/2021
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [Nearly 250 women have been fatally shot by police since 2015] (https://www.washingtonpost.com/graphics/2020/investigations/police-shootings-women/)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
#turn off scientific notation
options(scipen=999)
#load tidyverse
library(tidyverse)
#load lubridate 
library(lubridate)
#load refinr
library(refinr)

```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data
# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

fatal_police_shootings<- read_csv("data/fatal-police-shootings-data.csv")

# Clean required data and prepare for analysis if needed. 

#counts rows to get total number of fatal shootings
total_fatal_shootings <- fatal_police_shootings %>% 
  nrow()
#filters and counts to get number of Black women fatally shot
total_black_women_shot <- fatal_police_shootings %>% filter(gender == "F" & race == "B") %>% nrow()
#filters and counts to get number of fatal shootings where race is known
race_known <- fatal_police_shootings %>% filter(!is.na(race)) %>% nrow() 

```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1 - Jenna

* **Sentence text**: [“Since 2015, Black women have accounted for less than 1 percent of the overall fatal shootings in cases where race was known. But within this small subset, Black women, who are 13 percent of the female population, account for 20 percent of the women shot and killed and 28 percent of the unarmed deaths.”]
* **Analysis summary**: I was able to confirm the finding for the first part of the sentence, which stated that black women accounted for less than 1 percent of overall fatal shootings. However, when I calculated the remaining two parts of the sentence, I couldn't confirm the finding. The sentence states that black women account for 20% of women shot, but I got 18%. The more troubling of the two not confirmed was the last part. My work found that 12.6% of all unarmed deaths were black women, but the sentence states they account for 28%. A possible explanation would be that they included the "undetermined" entries in the unnamed category as well. Another possible explanation is because the story was published in Sept of 2020, and the data base was most recently updated in Sept. 2021. The trends in the data are relatively similar, but I believe that's why I cannot replicate their findings exactly.


```{r}
# Put code to reverse engineer sentence here
#(part 1)Black women have accounted for less than 1 percent of the overall fatal shootings

total_black_women_shot_percentage = (total_black_women_shot/race_known)*100
total_black_women_shot_percentage

# Display results of code below this codeblock
```

```{r}
# Put code to reverse engineer sentence here

# (part 2) Black women, who are 13 percent of the female population, account for 20 percent of the women shot
total_women_shot = fatal_police_shootings %>% 
   #total women shot
  filter(gender == "F") %>% nrow()
  #divide black women by total women shot
  black_women_out_of_women_shot = (total_black_women_shot/total_women_shot)*100
# Display results of code below this codeblock
   black_women_out_of_women_shot
```

```{r}
# Put code to reverse engineer sentence here

# (part 3) Black women account for 28 percent of the unarmed deaths.
total_unarmed = fatal_police_shootings %>% 
   #total women shot
  filter(armed == "unarmed") %>% nrow()
  #divide black women by total unarmed
  black_women_out_of_unarmed = (total_black_women_shot/total_unarmed)*100
# Display results of code below this codeblock
   black_women_out_of_unarmed 
```

### Sentence 2 - Jenna

* **Sentence text**: [“About 31 percent, or 77, of the 247 women fatally shot by police since 2015 had mental health issues, compared to 22 percent of the 5,362 men killed.”]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?] My entire analysis is a bit off from the data listed in the sentence. I believe that is because the story was published in Sept of 2020, and the data base was most recently updated in Sept. 2021. The trends in the data are relatively similar, but I believe that's why I cannot replicate their findings exactly. 

```{r}
# Put code to reverse engineer sentence here

#create new df of women shot
women_df <- fatal_police_shootings %>% 
  filter(gender=="F")
#count rows
  nrow(women_df)
  #294 women
#filter for signs_of_mental_illness
women_mental_illness<-  women_df %>% 
  filter(signs_of_mental_illness=="TRUE")
#count rows 
  nrow(women_mental_illness)
  #95 women with mental illness
#divide women by signs of mental illness and *100
(95/294)*100
  #32.3% 
    
#create new df of men shot
men_df <- fatal_police_shootings %>% 
  filter(gender=="M")
#count rows
  nrow(men_df)
  # 6303 men
#filter for signs_of_mental_illness
men_mental_illness<-  men_df %>% 
  filter(signs_of_mental_illness=="TRUE")
#count rows 
  nrow(men_mental_illness)
  #1399 men with mental illness
#divide men by signs of mental illness and *100
  (1399/6303)*100
  #22.2%
# Display results of code below this codeblock

```

### Sentence 3 - Taneen

* **Sentence text**: [“Crenshaw said Black women’s deaths also may be dismissed as “collateral damage” if they are killed while police are pursuing someone else. Twenty of the 247 women were killed in that kind of situation, analysis shows. In 12 of those 20 shootings, police said the women killed were caught in crossfire or shot accidentally.”]
* **Analysis summary**: [I was sort of able to confirm the finding. The Washington Post didn't define what they meant as collateral. I think the columns they used were “armed” and “flee.” I was able to get 20 women from counting “unarmed” and “not fleeing/NA.”
I was not able to confirm the second part because I don't think it is in the data. I think the way the WaPo got that information is by seeing the 20 women killed in collateral damage in the data, and then request police reports/interviews to get that detailed information.]

```{r}
# Put code to reverse engineer sentence here
unarmed_women<-fatal_police_shootings %>% 
  filter (gender=="F" & armed=="unarmed")
#(part 1) 20/247 women were killed as collateral damage
unarmed_women%>%
  group_by(flee)%>%
   summarise(number = n()) 

#(part 2) 12/20 were shot in crossfire or accidentally shot
  #I do not think they got this from the data

# Display results of code below this codeblock


### Sentence 4 - Jamie

* **Sentence text**: [Since 2015, police have killed 26 unarmed women, including Taylor. Of those, 14 were White, seven were Black, four were Hispanic and one’s race was unknown. While about twice as many White women were shot and killed as Black women, White women account for five times the Black female population"]
* **Analysis summary**: [For the first part, my result was that police killed 28 unarmed women, not 26. The second part I found that there were 15 White women and 5 Hispanic women instead, hence 2 more people in the outcome. The differences in the result might be because the story was published in September 2020, but the data that was used was updated in September 2021, which might have added new information.]

```{r}
# Put code to reverse engineer sentence here
#Since 2015, police have killed 26 unarmed women.
fatal_police_shootings%>%
  filter(gender=="F", armed=="unarmed")

#14 were White, seven were Black, four were Hispanic and one’s race was unknown
white_unarmed_women<-fatal_police_shootings%>%
  filter(gender=="F", armed=="unarmed", race=="W")%>%
  summarise(count = n())

black_unarmed_women<-fatal_police_shootings%>%
  filter(gender=="F", armed=="unarmed", race=="B")%>%
  summarise(count = n())

hispanic_unarmed_women<-fatal_police_shootings%>%
  filter(gender=="F", armed=="unarmed", race=="H")%>%
  summarise(count = n())
  
# Display results of code below this codeblock

```



### Sentence 5 - Max

* **Sentence text**: [Paste in sentence to engineer here]
“Of the 89 women killed at residences where they lived or often stayed, 12 encounters began as did Taylor’s — with a warrant to conduct a search or make an arrest in an investigation.”
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]
I was able to confirm some findings by assume that they used column names. 

```{r}
# Put code to reverse engineer sentence here

warrants_residences <-fatal_police_shootings %>%
filter (gender == "F" & location == "Residence") %>%
summarise(count=n())


Warrants_residences %>%
group_by(warrant) %>%
 summarise(number=n())



# Display results of code below this codeblock

```

-30-


