---
title: "lab_02"
author: "derek willis"
date: "8/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse and Lubridate. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse and lubridate
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with West Virginia loans contained in the PPP loans data.  

**Task** Run the codeblock below to load the data.

```{r}
west_virginia_ppp <- read_rds("lab_02.rds")
```

## Examine Data

## Answer questions

**Q1** Create a new column that divides the loan amounts into multiple options: under_10k, 10k_50k, 50k_100k, 100k_500k, 500k_1m, 1m_plus. How many loans are in each group?
**A1** 
10k_50k     15403			
under_10k	  9310			
100k_500k	  4132			
50k_100k    3968			
500k_1m	    539			
1m_plus	    356
```{r}
#select df you want to use
west_virginia_ppp %>%
  # chose columns you want to see in the table output
  select(loan_number, amount) %>%
  # creates a new column with all of these labels - be sure to enter numbers with no quotes and   characters with 'character here'
  mutate(
    loan_amount = case_when(
        amount < 10000 ~ "under_10k",
        amount >= 10000 & amount<=  50000 ~ "10k_50k",
        amount > 50000  & amount<= 100000 ~ "50k_100k",
        amount > 100000 & amount<= 500000 ~ "100k_500k",
        amount > 500000 & amount<=1000000 ~ "500k_1m",
        amount > 1000000 ~ "1m_plus"
      )
  ) %>% 
  # groups by labels
  group_by(loan_amount) %>% 
  # counts number each label
  summarise(number_of_loans=n()) %>%
  arrange(desc(number_of_loans))

```

**Q2** Which servicing lenders have the most loans with undisbursed amounts greater than 0?
**A2** Harvest Small Business Finance, LLC with 139 undisbursed amounts greater than 0.

```{r}
#select data frame
west_virginia_ppp %>% 
  #only show undisbursed amount greater than 0
  filter(undisbursed_amount > 0.00)%>% 
  #group these undisbursed loans by the service lender's names
  group_by(servicing_lender_name) %>% 
  #count each undisbursed loan for every lender
  summarize(total=n())%>%
  #display undisbursed loan count in discending order
  arrange(desc(total))
   
```

**Q3** How would you describe the top 10 list of those lenders? What do they have in common or not? (You may want to look up some of the top ones.)
**A3** Some of the top companies are mortgage/homeowner related companies or financial/banking companies. 

**Q4** Which month in 2021 had the most loans approved?
**A4** March 2021 with 4,404 -- I still need to figure out how to filter just for the 2021 year.

```{r}
west_virginia_ppp %>%
  mutate(Month = floor_date(date_approved, "month")) %>%
  group_by(Month) %>%
  summarise(total = n()) %>%
  arrange(desc(total))
```
