---
title: "lab_07"
author: "sean mussenden"
date: "8/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings
**Task**: Load rvest, janitor and the tidyverse
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse,rvest,janitor
library(rvest)
library(tidyverse)
library(janitor)
```


Q1. How many individual cases has the U.S. Department of Justice filed against people accused of fraud related to the PPP or EIDL loan program, according to the DOJ website listing those cases: https://www.justice.gov/criminal-fraud/cares-act-fraud?  An example of one case that appears on this page is "U.S. v. Richard Ayvazyan et al". To answer this question, you will need to use rvest scrape the data on this single webpage into a dataframe that lists only case names. Hint: you will need to make use of html_elements() and html_text() -- a function that extracts text inside of an html tag -- for this.
A1. 89 individual cases.

```{r}
cases_url <- "https://www.justice.gov/criminal-fraud/cares-act-fraud"

case_list <- cases_url %>%
  read_html() 
  
ul<- case_list %>%
  html_elements("ul")

case_names<- ul %>% html_elements("b") %>% 
  html_text() 

data.frame(case_names)

```

Q2. In how many individual judicial districts has the U.S. Department of Justice filed cases against people accused of fraud related to the PPP or EIDL loan program, according to the DOJ website listing those cases: https://www.justice.gov/criminal-fraud/cares-act-fraud?  Note: an example of a judicial district is "Southern District of Florida". You will need to use rvest scrape the data on this single webpage into a dataframe.
A2.28 individual judicial districts.

```{r}
p<- case_list %>%
  html_elements("p")

judicial_districts<- p %>% html_elements("i") %>% 
  html_text() %>% 
  data.frame()
```

Q4. The website (https://www.justice.gov/criminal-fraud/cares-act-fraud) shows that the DOJ filed more cases in the Southern District of Florida than in any other district. One of those cases was filed against someone named "Diamond Blue Smith". Who is Smith, and what was he accused of, according to the criminal complaint? If you were an editor, would you have assigned a reporter to write a story about this case when the complaint was filed in court? Why or why not?
A4. Diamond Blue Smith is a Florida rapper who is accused of wire fraud, bank fraud and attempt and conspiracy to commit wire and bank fraud. If I were a reporter I would assign a reporter to this story to write about the filing and accusations made against the rapper. I would also make sure they follow up and track the case and see what information comes out, if he ended up being charged, etc. 

Q5. In what percentage of all judicial districts has the U.S. Department of Justice filed cases cases against people accused of fraud related to the PPP or EIDL loan program, according to the DOJ website listing those cases: https://www.justice.gov/criminal-fraud/cares-act-fraud? In answering this question, you should also produce a list of judicial districts where DOJ has NOT filed a case, according to this site.  Note: to answer this question, you will need to scrape a table of all district courts on this up-to-date Wikipedia page under the heading "Active Courts": https://en.wikipedia.org/wiki/List_of_United_States_district_and_territorial_courts  
A5. The U.S. Department of Justice filed cases against 29.78% of all judicial districts.

```{r}
all_districts_url <- "https://en.wikipedia.org/wiki/List_of_United_States_district_and_territorial_courts"

all_districts <- all_districts_url%>%
  read_html() %>% 
  html_table()

all_districts  <- all_districts [[3]]

all_districts_count<- all_districts %>% nrow()

judicial_district_count <- judicial_districts %>% nrow()


(judicial_district_count/all_districts_count)*100



```
Q6. What might explain why, according to this site, the DOJ has filed PPP-related fraud cases in less than half of the country's judicial districts?
A6. Several states have multiple districts within their state (ie. there's a Northern, Middle and Southern District of Alabama). Just because less than half of the districts have cases filed against them doesn't mean that less than half the states themselves have cases filed against them. 

Q7. Which state had the most approved PPP loans per 100,000 population? [This web page](https://smussenden.github.io/coursefiles/ppp_scraping_example/index.html) has links to 52 individual web pages, one for each state (plus Washington, D.C. and Puerto Rico). Each of those web pages contains a one-row html table that has the name of the state, the number of approved loans, and the 2019 population. Here's an example for [Alabama](https://smussenden.github.io/coursefiles/ppp_scraping_example/states/alabama.html). You'll need to loop over the individual state urls, scrape each individual page and combine the information on each page into a single dataframe to answer this question.  
A7. South Dakota had the most approved PPP loans with 7379.531 loans per 100,000.
```{r}
#grab page that has urls
state_urls <- "https://smussenden.github.io/coursefiles/ppp_scraping_example/index.html"
#read the html and pull out the table that holds the urls within the html
state_url_list <- state_urls %>%
  read_html() %>% 
  html_table()
#only keep the section of html table that has the urls
state_url_list <- state_url_list [[1]]
#empty df to eventually put info
all_state_info <- tibble()
#loop through each url from state_url_list
for(row_number in 1:nrow(state_url_list)){
  #keep only the row for a given row number, get rid of every other row (so it doesn't keep repeating on the same row)
  each_row_df <- state_url_list %>%
      slice(row_number)
  #define url of page to get
  target_url <- each_row_df$url

#print(target_url)
  #create df that visits url from each state, reads inn table and extracts nested list
  state_info <- target_url %>%
    read_html() %>%
    html_table()
  #print each table (could comment this out eventually, but is a good way to check everythng is correct)
  print(state_info)
  3#put every row from each state into the empty df we created 
  all_state_info <- all_state_info %>%
      bind_rows(state_info)
  
}

all_state_info<- all_state_info %>% 
   mutate(
    pop_100000 = population/100000, 
    per_100000 = (total_ppp_loans/pop_100000)
  ) %>% 
  arrange(desc(per_100000))

```
