---
title: "lab_04"
author: "derek willis"
date: "8/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse.
library(tidyverse)
```

## Load Data

Let's keep working with Maryland loan applications contained in the PPP data.

**Task** Read in the lab_04.rds file and assign it to the variable `maryland_ppp`.

```{r}
maryland_ppp <- read_rds("lab_04.rds")
```

To answer the questions below, you'll need to join both the NAICS code data and the ZCTA demographic data from the book chapter. There's a particular set of steps you'll need to perform to make that work properly.

First, let's load the NAICS code CSV file we have in the data folder:

**Task** Read in the naics_codes.csv file and assign it to the variable `naics_codes`.

```{r}
naics_codes<- read_csv("naics_codes.csv")
```

Next join the two and then create a 5-character `zip5` column that we can use to join to the ZCTA data:

**Task** Write code to join the maryland_ppp and naics_codes data, assigning the result to a new variable. Then add a new `zip5` column.

```{r}
maryland_ppp %>% left_join(naics_codes) %>% select(name, naics_code, title)

maryland_ppp_with_naics_codes<- maryland_ppp %>% 
  left_join(naics_codes) %>% 
  mutate("zip5")
```

Let's add the Zip Code Tabulation Area dataset. You'll need to make the `ZCTA5N` column a character column, not a numeric one:

**Task** Read in the Zip Code Tabulation Area dataset in the data folder and change the datatype of the `ZCTA5N` column to character.

```{r}
maryland_zcta<- read.csv("maryland_zcta.csv")

maryland_zcta <- maryland_zcta %>% mutate(across(ZCTA5N, as.character))
```

Now we can join the PPP data to the ZCTA data:

**Task** Join the Maryland PPP data (with NAICS already joined) with the ZCTA data, specifying the join column in both dataframes.

```{r}
maryland_ppp_with_naics_codes <- maryland_ppp_with_naics_codes %>% mutate(zip5 = str_sub(zip, 1, 5))
maryland_ppp_with_naics_and_demographics <- maryland_ppp_with_naics_codes %>% left_join(maryland_zcta, by=c("zip5"="ZCTA5N"))
```

## Answer questions

**Q1** What are the top 5 industries (using the NAICS title) with the most PPP applications in Maryland?
**A1** 
Taxi Service	
Beauty Salons	
Full-Service Restaurants	
Offices of Real Estate Agents and Brokers	
All Other Personal Services

```{r}
maryland_ppp_with_naics_and_demographics %>%
  select(title) %>%
  group_by(title) %>% 
  summarise(count=n()) %>% 
  arrange(desc(count))
```

**Q2** What are the top industries (using the NAICS title) with the most PPP applications in Maryland where the loan amount is more than $150,000?
**A2** 	
Full-Service Restaurants	
Offices of Physicians (except Mental Health Specialists)	
Plumbing, Heating, and Air-Conditioning Contractors	
Offices of Dentists	
Commercial and Institutional Building Construction

```{r}
maryland_ppp_with_naics_and_demographics %>%
  select(title,amount) %>%
  filter(amount>150000) %>% 
  group_by(title) %>% 
  summarise(count=n()) %>% 
  arrange(desc(count))

```

**Q3** Which Maryland zip code with at least 20 percent non-Hispanic Asian population had the largest sum of PPP loans? Also show the number (count) of loans.
**A3** The zip code 20850 had the largest sum of loans at 474,525,057.53. This total comes from 3186	separate loans.

```{r}
maryland_ppp_with_naics_and_demographics %>%
  filter(NHA > 20) %>% 
  group_by(zip5) %>% 
  summarise(
    total_loan_amount= sum(amount),
    count=n()
    ) %>% 
  arrange(desc(count))

```

**Q4** Which Maryland zip code had the largest sum of loans where the reported race of the applicant was "Asian", and what is the percentage of non-Hispanic Asian population in that zip code? Also show the number (count) of loans.
**A4** The zip code 20850	also had the largest sum of loans, with 240 loans totaling 28,994,798.00, where the reported race of the applicant was "Asian." The percentage of non-Hispanic Asian population in this zip code is 23.5%.

```{r}
maryland_ppp_with_naics_and_demographics %>%
  filter(race == "Asian") %>% 
  group_by(zip5, PNHA) %>% 
  summarise(
    total_loan_amount= sum(amount),
    count=n()
    ) %>% 
  arrange(desc(count))

```

**Q5** What might explain why 20850 has a significant Asian population (23.5%) and only 240 of the 3,186 loans (7.5%) are to Asian applicants?
**A5** There is a large number of "unanswered" entries in the race column of the maryland_ppp data. There is a chance that there are more Asian owned businesses in the data then what is explicitly listed in the dataframe. 


**Q6** Find any Maryland zip codes that have loans where the PPP data lists the applicants as being Asian but where the zip code does not appear to have Asian population in the ZCTA data? What do you think is going on here?
**A6** Since ZCTAs are generalized information, it makes sense that there are a handful of loans in zip codes where ZCTA data lists NHA as 0.

```{r}
maryland_ppp_with_naics_and_demographics %>%
  filter(race == "Asian" & NHA == 0) %>% 
  group_by(zip5, NHA, race) %>% 
  summarise(
    count=n()
    ) %>% 
  arrange(desc(count))

```

**Q7** How many industries have exactly one loan in the data (hint: you can filter _after_ using `summarise` on any totals you create). From a newsworthiness standpoint, which of these is most worthy of further exploration, and why?
**A7** I think the two industries that are worth further exploration are the Ammunition (except Small Arms) Manufacturing and Small Arms Ammunition Manufacturing. As gun violence is an extremely salient topic, I think it's worth looking into the single applicants in each of these industries. And arguabley more importantly, why is there only one applicant per industry? Did ammunition companies not lose any business over the pandemic? Did their business get better? All of these questions are worth exploring based off of the information shown in the data. 

```{r}
maryland_ppp_with_naics_and_demographics %>%
  group_by(title) %>% 
  summarise(count=n()) %>% 
  filter(count==1)
```
